### 2016-07-06 filters-results.tex

\section{Achieving FP-Security and Privacy}
\tsnote{Under our current notions, you can never hope to prove privacy bounds that beat the min-entropy of $\distr{U}{n}$.}
\begin{itemize}
\item Prove privacy for the bigram/PRF-based double-hashing construction that Neidermayer et al.\ attacks. I'm pretty sure the attack is in the public-BB setting, and the provable security bound is considerably worse than the min-entropy because of the way the representation is made.  (Essentially, the Hamming weight of the representation is a good estimate of the length of the longest surname in filter.)  Note that their attack would allow one to recover \emph{every} name, not just one.

\item Prove that the basic BF, in the ROM, is private.  This also shows that the basic BF with $h_j(x)=F_K(\langle j,x \rangle)$ is private.  \tsnote{I believe this also shows that the bigram approach examined by Neidermayer et al.\ is secure when the hash functions are the $h_j(x)$ just described. }

\item Prove privacy of other suggestions, such as the record-level BF from Durham et al.?  This would be a nice pairing with the Neidermayer et al.\ results

\item Attack privacy of Dong, Chen, Wen construction in the BB-hash setting \tsnote{I give the attack in the related work section.  Copy it here and flesh it out.}  Prove privacy of construction in secret-hash setting

\item \tsnote{...}
\end{itemize}