### 2016-07-06 filters-results.tex
cheers

\section{Achieving FP-Security and Privacy}
\tsnote{Under our current notions, you can never hope to prove privacy bounds that beat the min-entropy of $\distr{U}{n}$.}
\begin{itemize}
\item Prove that the basic BF, in the ROM, is private.  \tsnote{Not at all surprising, but it is the first result one would want in this area.}

\item As a simple corrolary, show that the basic BF with $h_j(x)=F_K(\langle j,x \rangle)$ is private.  \tsnote{I believe this also shows that the bigram/PRF approach examined by Neidermayer et al.\ is secure when the hash functions are the $h_j(x)$ just described. }

\item Cast Neidermayer et al.\ attacks on the bigram/PRF-based double-hashing construction (see their paper) into our formalism.  \tsnote{Note that their attack would allow one to recover \emph{every} name, not just one.} \tsnote{Note that Neidermayer et al.\ suggest ditching the double-hashing construction in favor of independent PRFs, so the previous item basically explores their suggestion.}

\item Prove a privacy bound for the bigram/PRF-based double-hashing construction that Neidermayer et al.\ attacks. I'm pretty sure the attack is in the public-BB setting, and the provable security bound is considerably worse than the min-entropy because of the way the representation is made.  (Essentially, the Hamming weight of the representation is a good estimate of the length of the longest surname in filter.)

\item Attack privacy of Dong, Chen, Wen ``garbled Bloom Filter'' construction in the BB-hash setting \tsnote{I give the attack in the related work section.  Copy it here and flesh it out.}  Prove privacy of construction in secret-hash setting.  \tsnote{Makes the point that the application setting really matters.}

\item Prove privacy bounds of other suggestions, such as the record-level BF from Durham et al.?  This would be a nice pairing with the Neidermayer et al.\ results

\item Show implications and/or separations between FP and privacy notions.

\item \tsnote{...}
\end{itemize}