### 2016-07-12 Proofs.tex
Proof Theorem2

\section{Proofs}
\tsnote{Jon: ignore these for now.  This is Animesh's workspace.}
\begin{theorem}\label{thm1}
Fix $k, m, n > 0$ and define a hash-based filter as in section \ref{sec:hbf}, except the hash functions are defined by $h_j(x) =  F_{K_0}(x) + j\cdot F_{K_1}(x) \bmod m$, where $F \colon \calK \times U \sends [m]$ is a PRF. Let $Func(U,[m])$ be the set of all functions $\rho \colon \calK \times U \sends [m]$. Let $B_1, B_2$ be 2 adversaries attacking the PRF advantage of $F$, and $A$ be another adversary attacking the false positive game of hash-based filters with private hash functions, then 
\begin{equation}
\AdvFPSecHash{B,t}{A} \leq  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2} + q \cdot \epsilon
\end{equation}
Note: Here, $q$ is the number of queries made to the $\QryOracle$ oracle and $\epsilon$ is the false positive error in the double hashing scheme given by Kirsch and Mitzenmacher. $\epsilon \rightarrow 0$, if $m/n \rightarrow \infty$ and/or $k \rightarrow \infty$ \cite{xxx}
\end{theorem}

\begin{proof}
The proof of the theorem will use a game playing argument as shown in Fig. \ref{fig:Game}. $\game{0}(A)$ exactly simulates $\ExpFPSecHash{B,t}{A}$ with hash functions $h_j(x) =  F_{K_0}(x) + j\cdot F_{K_1}(x) \bmod m$. So,
\begin{equation}
\AdvFPSecHash{B,t}{A} = \Prob{\game{0}(A)=1}\label{eq:0}
\end{equation}
 $\game{1}(A)$ and $\game{2}(A)$ attack 2 different filters with hash functions 
\begin{eqnarray*}
h_j(x) &=& ( F_{K_0}(x) + j\cdot \rho_2(x)) \bmod m\\
\mathrm{and,}\nudge h_j(x) &=& ( \rho_1(x) + j \cdot \rho_2(x)) \bmod m 
\end{eqnarray*}

Here, $\rho_1$ and $\rho_2$ are random functions sampled from $Func(U,[m])$. Let adversaries $B_1, B_2$ simulate $\game{1}(A)$ and $\game{2}(A)$ respectively as shown in Fig. \ref{fig:BGame1}. When $b=1$, $B_1$ and $B_2$ exactly simulate $\game{0}(A)$
and $\game{1}(A)$. While, for $b=0$, $B_1$ wins, if $\game{1}(A)$ outputs 0, and  $B_2$ wins, if $\game{2}(A)$ returns 0.

\begin{eqnarray}
\Prob{\ExpPRF{F}{B_1} = 1|b=1} &=& \Prob{\game{0}(A)=1}\label{eq:m1}\\
\Prob{\ExpPRF{F}{B_2} = 1|b=1} &=& \Prob{\game{1}(A)=1}\label{eq:m2}\\
\Prob{\ExpPRF{F}{B_1} = 1|b=0} &=& 1-\Prob{\game{1}(A)=1}\label{eq:m3}\\
\Prob{\ExpPRF{F}{B_2} = 1|b=0} &=& 1-\Prob{\game{2}(A)=1}\label{eq:m4}
\end{eqnarray}

\par{In a classical Bloom filter, adversary $A^{'}$ is non-adaptive, i.e., for some fixed $x \notin S$, $A^{'}$ targets the false-positive error conditioned on randomness of hash functions. But in the $\ExpFPSecHash{B,t}{A}$ game, $A$ is adaptive, as it can adapt its queries as well as its output based on response of $\QryOracle$ oracle. Let $A$ makes $q>t$ queries $x_1, x_2, ..., x_q$ to the $\QryOracle$ oracle. $\game{3}(A)$ simulates $fp-error$ game of $A^{'}$ as shown in Fig. \ref{fig:Game}.}

%\caption{Game playing argument}\label{fig:Game}
\begin{figure}
\fpage{.95}{
\hpagessd{.5}{.5}
{
\underline{\game{0}(A)}\\
$K_1, K_2 \getsr \calK$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_i(x) = F_{K_1}(x)+i \cdot F_{K_2}(x)$, for $i\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
{
\underline{\game{1}(A)}\\
$K_1 \getsr \calK$\\
$\rho \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_i(x) = F_{K_1}(x)+i \cdot \rho(x)$, for $i\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
{
\underline{\game{2}(A)}\\
$\rho_1,\rho_2 \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_i(x) = \rho_1(x)+i \cdot \rho_2(x)$, for $i\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
{
\underline{$\game{3}(A)$}\\
$\ell \getsr [q]$, where $[q] = {0,1, \cdots, q-1}$\\
$count = 0$ \\%count will be used to keep track of queries made to $\QryOracle$\\
$win \gets false$\\
$\rho_1,\rho_2 \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
$count=count +1$\\
if $(count == \ell)$\\
$\nudge win = true$\\
$\nudge \mathrm{Ret} \nudge \Qry^{\HashOracle}(M,\tau,x)$\\
else \\
$\nudge \mathrm{Ret}\nudge 0$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_i(x) = \rho_1(x)+i \cdot \rho_2(x)$, for $i\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
}
\caption{Theorem \ref{thm1}:Game playing argument}\label{fig:Game}
\end{figure}

%\caption{Adversary $B_1, B_2$ simulating $\game{1}(A)$ and $\game{2}(A)$ respectively}\label{fig:BGame1}
\begin{figure}
\centering
\fpage{0.9}{
\hpagessl{0.5}{0.5}
{
$\experimentv{Adversary \nudge B_1^{\calO}:}$\\
$K_1 \getsr \calK $\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
When $A$ asks $\QryOracle(x)$:\\
$\nudge$ Ret $\Qry^{\HashOracle}(M,\tau,x)$\\
When $A$ halts with output $\calX^{'}$: \\
if $|\calX^{'}| < t$ or $\calX^{'} \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX^{'}$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
\oracle{$\HashOracle(x)$}\\
$ \mathrm{Ret} \nudge F_{K_1}(x) + i\cdot \calO(x)$, for $i\in[1,k]$\\
}
{
$\experimentv{Adversary \nudge B_2^{\calO}:}$\\
%$d \getsr \bits $\\
$\rho_2 \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
When $A$ asks $\QryOracle(x)$:\\
$\nudge$ Ret $\Qry^{\HashOracle}(M,\tau,x)$\\
When $A$ halts with output $\calX^{'}$: \\
if $|\calX^{'}| < t$ or $\calX^{'} \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX^{'}$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
\oracle{$\HashOracle(x)$}\\
%if $d=0$\\
$\mathrm{Ret} \nudge \calO(x) + i\cdot \rho_2(x)$, for $i\in[1,k]$
}
}
\caption{Theorem \ref{thm1}:Adversary $B_1, B_2$ simulating $\game{1}(A)$ and $\game{2}(A)$ respectively}\label{fig:BGame1}
\end{figure}

From eqns. \ref{eq:m1}, \ref{eq:m2}, \ref{eq:m3} and \ref{eq:m4},
\begin{eqnarray}
\nonumber \Prob{\ExpPRF{F}{B_1} = 1} &=& .5\Prob{\ExpPRF{F}{B_1}=1|b=0} + .5\Prob{\ExpPRF{F}{B_1} = 1|b=1}\\
\nonumber 2\Prob{\ExpPRF{F}{B_1} = 1} &=& 1-\Prob{\game{1}(A)=1} + \Prob{\game{0}(A)=1}\\
\nonumber (2\Prob{\ExpPRF{F}{B_1} = 1} - 1)  &=& \Prob{\game{0}(A)=1} - \Prob{\game{1}(A)=1}\\
 \AdvPRF{F}{B_1} &=& \Prob{\game{0}(A)=1} - \Prob{\game{1}(A)=1} \label{eq:1}\\
\nonumber& &\\
\nonumber \Prob{\ExpPRF{F}{B_2} = 1} &=& .5\Prob{\ExpPRF{F}{B_2}=1|b=0} + .5\Prob{\ExpPRF{F}{B_2} = 1|b=1}\\
\nonumber 2\Prob{\ExpPRF{F}{B_2} = 1} &=& 1-\Prob{\game{2}(A)=1} + \Prob{\game{1}(A)=1}\\
\nonumber (2\Prob{\ExpPRF{F}{B_2} = 1} - 1)  &=& \Prob{\game{1}(A)=1} - \Prob{\game{2}(A)=1}\\
 \AdvPRF{F}{B_2} &=& \Prob{\game{1}(A)=1} - \Prob{\game{2}(A)=1} \label{eq:2}
\end{eqnarray}

Adding eqn. \ref{eq:1} and \ref{eq:2}, we have
\begin{equation}
\AdvPRF{F}{B_2} +\AdvPRF{F}{B_1} = \Prob{\game{0}(A)=1} - \Prob{\game{2}(A)=1}\label{eq:2a}
\end{equation}

In $\game{3}(A)$, assume, without loss of generality that $A$ halts with output $x$ immediately, if it asks $x$ to $\QryOracle$ and $\QryOracle$ returns $1$. Thus, all queries to $\QryOracle$ return 0 until the first one. Hence,

\begin{equation}
\Prob{\game{2}(A)=1} \leq q \cdot \Prob{\game{3}(A)=1} = q \cdot \Prob{\game{3}(A)=1|win=true}\label{eq:3}
\end{equation}

In Fig. \ref{AGame3}, adversary $A^{'}$ simulates $\game{3}(A)$. Since, adversary $A^{'}$ exactly simulates $\game{3}(A)$ conditioned on $win = true$, 
\begin{eqnarray}
\nonumber \Prob{\game{3}(A)|win=true} &=& \Prob{A^{'} \mathrm{outputs \nudge FP}}\\
& \leq & \epsilon, \nudge \mathrm{where} \nudge \epsilon = (1-e^{-kn/m})^k\label{eq:4}
\end{eqnarray}

%\caption{Adversary $A^{'}$ simulating $\game{3}(A)$} \label{AGame3}
\begin{figure}
\centering
\fpage{0.4}{
$\experimentv{Adversary \nudge A^{{'}\HashOracle}:}$\\
$\ell \getsr [q]$, where $[q] = {0,1, \cdots, q-1}$\\
$count = 0$\\
$\rho_1,\rho_2 \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
When $A$ asks $\QryOracle(x)$:\\
$\nudge cnt = cnt + 1$\\
\nudge if $count == \ell$\\
$\nudge$ $\nudge r \gets \Qry^{\HashOracle}(M,\tau,x)$\\
\nudge else\\
$\nudge$ $\nudge r = 0$\\
\nudge Ret $r$\\
When $A$ halts with output $\calX^{'}$\\
\nudge if $|\calX^{'}| < t$ or $\calX^{'} \cap \calS \neq \emptyset$ then \\
\nudge\nudge Ret 0\\
\nudge if $\exists x \in \calX^{'}$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge \nudge Ret 0\\
\nudge Ret 1\\
%
\oracle{$\HashOracle(x)$}\\
$h_i(x) = \rho_1(x)+i \cdot \rho_2(x)$, for $i\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
\caption{Theorem \ref{thm1}:Adversary $A^{'}$ simulating $\game{3}(A)$} \label{AGame3}
\end{figure}

From eqn. \ref{eq:0}, \ref{eq:2a}, \ref{eq:3} and \ref{eq:4}: \\
\begin{eqnarray}
\nonumber \AdvFPSecHash{B,t}{A} &-& \Prob{\game{2}(A)=1} =  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}   \\
\nonumber \AdvFPSecHash{B,t}{A}  &=&  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2} + \Prob{\game{2}(A)=1}  \\
\AdvFPSecHash{B,t}{A} &\leq&  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}  + q \cdot \epsilon
\end{eqnarray}
\end{proof}

\newcommand{\FK}{F_K(\langle j,x \rangle)}
\newcommand{\rhoK}{\rho(\langle j,x \rangle)}
\begin{theorem}\label{thm2}
A hash-based filter with private hash functions $h_j(x) = \FK$ is secure in the false positive game of an adaptive adversary (A) who makes $q$ set-membership queries in time $O(t)$ , if F is PRF secure and the filter has low FP-error($\epsilon$) in the non-adaptive setting. Here, adversary $B$ attacks the PRF game in time $O(t)$, makes $q$ oracle queries, and 
\begin{equation}
\AdvFPSecHash{B,t}{A} \leq  \AdvPRF{F}{B} + q \cdot \epsilon
\end{equation}
Note: $\FK$ provides domain separation and hence mutual independence, unlike hash functions in theorem \ref{thm1}
\end{theorem}

\acnote{Not complete yet}
\begin{proof}
We use a game playing argument similar to theorem \ref{thm1} to prove this theorem. In \game{0}(A), adversary attacks the FP game with private hash functions $h_j(x) = \FK $, and hence exactly simulates $\ExpFPSecHash{B,t}{A}$.
\begin{equation}
\AdvFPSecHash{B,t}{A} = \Prob{\game{0}(A)=1}\label{eq:0T2}
\end{equation}
\game{1}(A) simulates $\ExpFPSecHash{B,t}{A}$ with a random function $\rho \getsr Func(U,[m])$ used to generate uniformly random  hash-indices with domain separation, of elements $x \in U$. Let $A^{'}$ be the non-adaptive friend of $A$. \game{2}(A) simulates behavior of $A^{'}$ as closely as possible. The three games are shown in Fig. \ref{fig:GameT2}.

%\caption{Game playing argument}\label{fig:Game}
\begin{figure}
\fpage{.93}{
\hpagesssl{.3}{.3}{.3}
{
\underline{\game{0}(A)}\\
$K \getsr \calK $\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_j(x) = \FK$, for $j\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
{
\underline{\game{1}(A)}\\
$\rho \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_j(x) = \rhoK $, for $j\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
{
\underline{$\game{2}(A)$}\\
$\ell \getsr [q]$, where $[q] = {0,1, \cdots, q-1}$\\
$count = 0$ \\%count will be used to keep track of queries made to $\QryOracle$\\
$win \gets false$\\
$\rho \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$\calX \getsr A^{\QryOracle}(\calS)$\\
if $|\calX| < t$ or $\calX \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
%
\oracle{$\QryOracle(x)$}\\
$count=count +1$\\
if $(count == \ell)$\\
$\nudge win = true$\\
$\nudge \mathrm{Ret} \nudge \Qry^{\HashOracle}(M,\tau,x)$\\
else \\
$\nudge \mathrm{Ret}\nudge 0$\\\\
%
\oracle{$\HashOracle(x)$}\\
$h_j(x) = \rhoK$, for $j\in[1,k]$\\
Ret $\left(h_1(x),\ldots,h_k(x)\right)$
}
}
\caption{Theorem \ref{thm2}:Game playing argument}\label{fig:GameT2}
\end{figure}

In Fig. \ref{fig:BGame1T2}, adversary $B$ simulates $\game{1}(A)$ and $\game{2}(A)$ for $b=1$ and $b=0$ respectively .

%\caption{Adversary $B$ and $A^{'}$ simulating $\game{1}(A)$ and $\game{2}(A)$ respectively}
\begin{figure}
\centering
\fpage{0.9}{
\hpagessl{0.5}{0.5}
{
$\experimentv{Adversary \nudge B^{\calO}:}$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
When $A$ asks $\QryOracle(x)$:\\
$\nudge$ Ret $\Qry^{\HashOracle}(M,\tau,x)$\\
When $A$ halts with output $\calX^{'}$: \\
if $|\calX^{'}| < t$ or $\calX^{'} \cap \calS \neq \emptyset$ then \\
\nudge Ret 0\\
if $\exists x \in \calX^{'}$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge Ret 0\\
Ret 1\\\\
\oracle{$\HashOracle(x)$}\\
$ \mathrm{Ret} \nudge \calO(\langle j,x \rangle)$, for $j\in[1,k]$\\
}
{
$\experimentv{Adversary \nudge A^{'\HashOracle}:}$\\
$\ell \getsr [q]$, where $[q] = {0,1, \cdots, q-1}$\\
$count = 0$\\
$\rho \getsr Func(U,[m])$\\
$S \getsr \distr{U}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
When $A$ asks $\QryOracle(x)$:\\
$\nudge cnt = cnt + 1$\\
\nudge if $count == \ell$\\
$\nudge$ $\nudge r \gets \Qry^{\HashOracle}(M,\tau,x)$\\
\nudge else\\
$\nudge$ $\nudge r = 0$\\
\nudge Ret $r$\\
When $A$ halts with output $\calX^{'}$\\
\nudge if $|\calX^{'}| < t$ or $\calX^{'} \cap \calS \neq \emptyset$ then \\
\nudge\nudge Ret 0\\
\nudge if $\exists x \in \calX^{'}$ s.t. $\Qry^{\HashOracle}(M,\tau,x)=0$ then\\
\nudge \nudge Ret 0\\
\nudge Ret 1\\\\
%
\oracle{$\HashOracle(x)$}\\
$\mathrm{Ret} \nudge \rhoK$, for $j\in[1,k]$
}
}
\caption{Theorem \ref{thm2}:Adversary $B$ and $A^{'}$ simulating $\game{1}(A)$ and $\game{2}(A)$ respectively }\label{fig:BGame1T2}
\end{figure}

\begin{eqnarray}
\Prob{\ExpPRF{F}{B} = 1|b=1} &=& \Prob{\game{0}(A)=1}\label{eq:m1T2}\\
\Prob{\ExpPRF{F}{B} = 1|b=0} &=& 1-\Prob{\game{1}(A)=1}\label{eq:m2T2}
\end{eqnarray}

From eqns. \ref{eq:m1T2}, \ref{eq:m2T2},
\begin{eqnarray}
\nonumber \Prob{\ExpPRF{F}{B} = 1} &=& .5\Prob{\ExpPRF{F}{B_1}=1|b=0} + .5\Prob{\ExpPRF{F}{B_1} = 1|b=1}\\
\nonumber 2\Prob{\ExpPRF{F}{B} = 1} &=& 1-\Prob{\game{1}(A)=1} + \Prob{\game{0}(A)=1}\\
\nonumber (2\Prob{\ExpPRF{F}{B} = 1} - 1)  &=& \Prob{\game{0}(A)=1} - \Prob{\game{1}(A)=1}\\
 \AdvPRF{F}{B} &=& \Prob{\game{0}(A)=1} - \Prob{\game{1}(A)=1} \label{eq:1T2}
\end{eqnarray}

Following a similar argument from theorem \ref{thm1},
\begin{equation}
\Prob{\game{1}(A)=1} \leq q \cdot \Prob{\game{2}(A)=1} = q \cdot \Prob{\game{2}(A)=1|win=true}\label{eq:3T2}
\end{equation}
\begin{eqnarray}
\nonumber \Prob{\game{2}(A)|win=true} &=& \Prob{A^{'} \mathrm{outputs \nudge FP}}\\
& \leq & \epsilon, \nudge \mathrm{where} \nudge \epsilon = (1-e^{-kn/m})^k\label{eq:4T2}
\end{eqnarray}

From eqn. \ref{eq:0T2}, \ref{eq:1T2}, \ref{eq:3T2},  and \ref{eq:4T2}: \\
\begin{eqnarray}
\nonumber \AdvFPSecHash{B,t}{A} &-& \Prob{\game{1}(A)=1} =  \AdvPRF{F}{B}   \\
\nonumber \AdvFPSecHash{B,t}{A}  &=&  \AdvPRF{F}{B} + \Prob{\game{1}(A)=1}  \\
\AdvFPSecHash{B,t}{A} &\leq&  \AdvPRF{F}{B}  + q \cdot \epsilon
\end{eqnarray}

\end{proof}
\clearpage
\section{Summary of attacks(based on Bloom filter exploit) on Squid}
\acnote{For reference}\\
\heading{Crossby and Wallach: Squid Attack on hash tables} 
\begin{itemize}
\item A Squid server was ran on a stand-alone machine.
\item A file containing a list of URL's was prepared.
\item The URL requests were parsed and served by the Squid server. This completed the set-up phase.
\item Crossby and Wallach targeted hash collisions in Squid to increase the average URL load time. The authors quote ``This attack does not represent a ''smoking gun” for algorithmic complexity attacks, but it does illustrate how common network services may be sensitive to these attacks". 2 experiments were carried out to illustrate this.
\item In experiment 1, around 143K random URL's were requested to the proxy server, whereas in experiment 2, same number of carefully chosen URL's that collide with the already present URL's in the hash-table were requested. 
\item Observation: The load time of random URL's was less than the chosen URL's. It must be noted that, the authors did not use a brute-force approach to find collisions. Rather,  they proposed an efficient method to find hash-collisions. The method involves finding a set of elements that hash to the same index and using a combination of those elements to find more elements. 
\end{itemize}

\heading{Gerber et. al. : Squid attack on cache digest:}
\begin{itemize}
\item Squid servers follow a hierarchical model, and in this model the peers can exchange summary of their cache using Bloom filters(cache-digest). 
\item In the set-up phase, the authors configure 2 peer Squid servers for caching and an HTTP server which responds to all queries of the Squid servers. 
\item Peer 1's cache-digest is polluted using fake URL's. Both the peers exchange their cache-digests 
\item Client(s) of Peer 2, start making URL requests. The URL's which are not cached by the latter are searched in its neighbors cache-digest and false positives are counted.
\item Observation: There is a substantial increase in FP's due to pollution as compared to FP's without it.
\item Reason: Fake URL's increase the Hamming weight of the cache digest of peer 1 and hence increase false positives.
\end{itemize}