### 2016-07-25 privacy.tex
adding privacy experiments

\subsection{Notions of Privacy}
One can also consider the orthogonal notion of privacy for a set-representation.
\jnote{Actually, I am now unsure what the right way to define it is. It seems natural
to give the attacker $\aux$. But if we give them $\aux$ and $M$ then they have everything.
If $S$ has high min-entropy (element-wise), then it might still be
possible to have security here.}\tsnote{I think we will always bump up
against the (element-wise) min-entropy of the distribution over~$S$,
at least in the ``one-wayness'' style definition that I had been imagining. }
There are several possible definitions:
\begin{enumerate}
\item A ``one-wayness'' definition where $S$ has high min-entropy (element-wise), and we
give the attacker $\aux$ and either $M$ or oracle access to $\Qry$. The attacker
succeeds if it can output an element in~$S$.

\item An ``indistinguishability'' definition where either $S_0$ or $S_1$ is used, and the attacker
must determine which. Here the attacker is given $\aux$ but only oracle access to $\Qry$, and
it is disallowed from asking any query $q$ for which $q(S_0) \neq
q(S_1)$. \tsnote{How will the adversary know that a query satisfies
  this in (say) the case of the classical BF?  We can, of course,
enforce it in the experiment, but then our resource accounting risks
overcounting the number of queries.} \jnote{I was assuming the attacker knows $S_0, S_1$.}
\tsnote{Also, we need that, in general, $\aux$ itself
does not leak information about~$S_0$ vs.\ $S_1$. } \jnote{Sure, this would be implied
by the definition in this case.}
\end{enumerate}

\begin{figure}[htp]
\centering
\hfpages{.45}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpWPriv{\setprim,\distr{\calS}{}}{A}$}\\
$S \getsr \distr{\calS}{}$\\
$(M,\aux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(\aux)$\\
if $z \in S$ then Return 1\\
Return 0

\medskip
\experimentv{$\ExpPriv{\setprim,\distr{\calS}{}}{A}$}\\
$S \getsr \distr{\calS}{}$\\
$(M,\aux) \getsr \Rep(S)$\\
$z \getsr A(M,\aux)$\\
if $z \in S$ then Return 1\\
Return 0
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\aux,q)$\\
Return~$a$
}
}
%
{
\hpagess{.5}{.45}
{
\experimentv{$\ExpPriv{\setprim}{A}$}\\
$b \getsr \bits$\\
$z \getsr A^{\RepOracle\TestOracle}(\aux)$\\
if $z = b$ then Return 1\\
Return 0
}
%
{
\oracle{$\RepOracle(S_0,S_1)$}\\
if $\mathrm{used}=\true$ then \\
\nudge Ret $\bot$\\
$\mathrm{used}=\true$\\
$(M,\aux) \getsr \Rep(S_b)$\\
Return $\aux$\\

\medskip
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\aux,q)$\\
Return~$a$
}
}
\caption{Privacy notions. (Still playing...)} 
\label{fig:privacy}
\end{figure}