### 2016-09-05 Proofs.tex
Syntax consistence

\section{Proofs}
\input{Theorem1_FP}

\tsnote{Jon: ignore these for now.  This is Animesh's workspace.}
\input{Theorem2_FP}
\input{Theorem3_FP}