### 2016-09-30 Proofs.tex
moving files to better names

\section{Proofs}
\input{lin-correctness}

\tsnote{Jon: ignore these for now.  This is Animesh's workspace.}
\input{Theorem2_FP}
%\input{Theorem3_FP}