### 2016-09-30 Proofs.tex
moving files, moving proofs inline

\section{Proofs}
%\input{lin-correctness}

\tsnote{Jon: ignore these for now.  This is Animesh's workspace.}
\input{ds-correctness}
%\input{Theorem3_FP}