### 2016-09-30 Proofs.tex
cheers

\section{Proofs}
%\input{lin-correctness}

\tsnote{Jon: ignore these for now.  This is Animesh's workspace.}
%\input{ds-correctness}
%\input{Theorem3_FP}