### 2016-09-30 results.tex
citations

\section{Case Studies}
Here we explore various set-membership data structures that have been proposed
in the literature with respect to our definitions.

%%%%%%%%%%%%%%%%% CLASSICAL BLOOM FILTER IN ROM %%%%%%%%%%%%%%%%%%%%

\begin{figure}[tp]
\centering
\hfpagess{.4}{.4}
{
\algorithmv{$\Rep^H(S)$}\\
$M \gets 0^m$\\
for $x \in S$\\
\nudge for $j \in \{1,2,\ldots,k\}$\\
\nudge\nudge $h_j \gets H(j,x)$\\
\nudge\nudge $M[h_j] \gets 1$\\
$\pubaux \gets M$, $\privaux \gets \emptystring$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry^H(\pubaux,\privaux,q_x)$}\\
$M\gets\pubaux$\\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets H(j,x)$\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1
}
{
\algorithmv{$\Rep(S)$}\\
$M \gets 0^m$; $K \getsr \calK$\\
for $x \in S$\\
\nudge for $j \in \{1,2,\ldots,k\}$\\
\nudge\nudge $h_j \gets F_{K}(\langle j,x \rangle) $\\
\nudge\nudge $M[h_j] \gets 1$\\
$\pubaux \gets M$, $\privaux \gets \langle K \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$M\gets\pubaux$, $K \gets \privaux$\\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets F_{K}(\langle j,x \rangle) $\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1
}
\caption{{\bf Left:} $\Rep$ and $\Qry$ algorithms for classic Bloom
  filter $\setprim_{\mathrm{Bloom}}$ in the ROM for a hash function $H\colon \mathbb{N} \times
  \elts \to [m]$. {\bf Right:} $\Rep$ and $\Qry$
  algorithms for the domain-separated PRF data structure
  $\setprim_{\mathrm{ds}}$, using function family $F \colon
  \calK \times \elts \to [m]$.  Both constructions have query set is $\calQ=\{q_x\colon
  \univ \to \bits\}$ where $q_x(S)=1 \Leftrightarrow x \in S$.  }
\label{fig:bf-and-ds}
\end{figure}

\subsection{Classical Bloom Filter in the Random-Oracle Model}

Fix $n,k,m \geq 0$, and fix a set $\elts$.  Then the classical Bloom
filter set-membership data structure $\setprim_{\mathrm{Bloom}}=
(\Rep, \Qry)$ over~$\elts$ is defined as in
Figure~\ref{fig:bf-and-ds} (left side).  We
define~$\setprim_{\mathrm{Bloom}}$ in the random-oracle model
(ROM), as the traditional analysis of Bloom filter correctness (as a
function of $n,k,m$) is done in this model.\footnote{The standard
model version would have $\Rep$ choose $H_1,H_2,\ldots,H_k$ from a
hash function family, compute $h_j$ using these, and placing their
description in $\pubaux$.}

\heading{Correctness. } The following theorem proves that the Bloom
filter meets our correctness notion in the ROM.  Again, we stress
that this correctness is with respect to adaptive adversaries,
aiming to find $r \geq 1$ errors (in this case, false-positives).
The proof appears in Appendix~\ref{sec:thm3_fp}. \jnote{We seem to
have plenty of space, so maybe start moving things from the appendix
to the main body?}

\begin{theorem}\label{thm3}\label{thm:bf-correctness}
Fix $k,m,n,r>0$, a set~$\elts$, and let $\setprim_{\mathrm{Bloom}}= (\Rep, \Qry)$ be classical Bloom filter (set-membership) data structure over~$\elts$. For any adversary~$A$ that makes a total of~$q_T$ queries to the $\Test$ oracle, and $q_R$ queries to the RO, and has time-complexity~$O(t)$,
\[
\AdvCorrect{\setprim_{\mathrm{Bloom}},r}{A} \leq  {\dbinom{q_T + q_H}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
\end{theorem}
We note that Kirsch and Mitzenmacher~\cite{kirsch2006less} show
that the ($r=1$, zero-query) error-probability for~$\setprim_\mathrm{Bloom}$ is
always within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic
rate of convergence is faster than~$1/n$.  The multiplicative (``hybrid'') term arises as a result
of moving from the adaptive to the zero-query setting.
Finally, $\binom{q_T+q_H}{r} \leq (q_T+q_H)^r$ (with reasonable tightness when~$r$ is small) in
which case the final term in the bound behaves as
$((q_T+q_H) (1-e^{-kn/m})^k + O(q/n) )^r$; loosely, this is the bound one expects for
trying to find a single correctness error (in this case, a false positive) in each of~$r$ independent``rounds'', each round
consisting of~$(q_T+q_H)$ attempts.
\input{BF_correctness}

\heading{Privacy. }
Turning to privacy, we observe immediately that the classical BF, in the
ROM, does not meet our SS-privacy notion when the leakage function
fails to reveal the entire set~$S$.  This is because the set~$S$ is
adversarially chosen, and the adversary has access to the random
oracle, i.e. the hash functions used to make~$\pubaux$.  Thus the
adversary can simply check that the $\pubaux$ it receives is correct.
Thus, unless $m,k,n$ are such that the false-positive rate is large,
the simulator needs to call the random oracle on the points in~$S$ in order to ``fool'' the adversary.
%\tsnote{I'm still not completely comfortable with this.  It isn't explicit in our SS notion, but I assume that~$A$ is stateful and so knows~$S$ in its second stage?  If so, I wonder if the classic BF in the ROM would be secure relative to a weaker notion (but still stronger than OW) in which the second stage does not know~$S$?  This seems likely to me.}

We note that this will be true for \emph{any} data structure that returns an empty $\privaux$, because this implies that $\pubaux$ alone suffices to evaluate $\Qry(\pubaux,\privaux,\cdot)$.

\begin{theorem}[{\rm OW-privacy of classic BF in the ROM}]\label{thm3}\label{thm:bf-ow}
Fix $k,m,n,r>0$, a set~$\elts$, and let $\setprim_{\mathrm{Bloom}}= (\Rep, \Qry)$ be the classical Bloom filter data structure over~$\elts$. Let $\distr{}{}$ be a distribution over subsets of universe~$\univ=2^\elts$, where $\mu = H_\infty(\distr{}{})$.  For any adversary~$A$ that makes $q$~queries to its oracle
\[
\AdvPrivOW{\setprim_{\mathrm{Bloom}},\distr{}{},r}{A} \leq  \frac{(kqn+n)}{2^{\mu}}\;.
\]
\end{theorem}
\todo{Typeset simple proof.  Result is on chalkboard.}
\begin{proof}
\begin{figure}[tp]
\hfpagess{.45}{.45}
{\underline{\game{0}($A$)}\\
$S \getsr \distr{}{}$\\
$(\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A(\pubaux)$\\
if $z \in S$ then return 1\\
return 0\\

\medskip
\oracle{$\HashOracle(\qry_x)$}\\
}
{
}
\end{figure}
\end{proof}

%\jnote{Organize by scheme, not by security notions.}

%%%%%%%%%%%%%%%%% DOMAIN-SEPARATED PRF %%%%%%%%%%%%%%%%%%%%

\subsection{Domain-separated PRF-Hash}
%\heading{Construction of PRF with domain separation.}
\heading{Correctness. }
Let $\setprim_{\mathrm{ds}}= (\Rep, \Qry)$ be the set-membership data
structure defined on the right-side of Figure~\ref{fig:bf-and-ds}.
As with the preceding theorem, the following says (informally) that
if~$F$ is a secure PRF, then this domain-separated PRF construction is
correct against adaptive adversaries.
% \tsnote{Not really appropriate to say this next thing; I'll add a new section that addresses bigram DS-PRF hash.}
% We note that Neidermayer et
% al. in~\cite{xxx} suggest replacing the linear-PRF hash functions in
% the previous construction with independently keyed PRFs as a way to
% thwart their attacks.  The construction we consider now achieves the
% same end, albeit via explict domain separation rather than separate
% keys.

\begin{theorem}[{\rm Correctness of DS-PRF}]\label{thm2}\label{thm:ds-correctness}
Fix $k,m,n,r>0$, a set~$\elts$, and let $\setprim_{\mathrm{ds}}= (\Rep, \Qry)$ be the set-membership data structure (over~$\elts$) in Figure~\ref{fig:bf-and-ds}.  For any adversary~$A$ that makes a total of~$q$ queries to its oracle and has time-complexity~$O(t)$, there exists a prf-adversary ~$B$ (explicitly constructed in the proof of this theorem) such that
\[
\AdvCorrect{\setprim_{\mathrm{ds}},r}{A} \leq  \AdvPRF{F}{B}  + {\dbinom{q}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B$ asks $q$ queries, and has time complexity $O(t+qm)$.
\end{theorem}
\fixme{This result is pretty intutitive following the classic BF correctness result. Probably fine to just explain it and push an appendix, or even drop from camera ready if out of space.}
We note that Kirsch and Mitzenmacher~\cite{kirsch2006less} show
that the ($r=1$, zero-query) error-probability for~$\setprim_\mathrm{ds}$ is
always within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic
rate of convergence is faster than~$1/n$.   As before, the multiplicative term $\binom{q}{r}$ arises when switching from the adaptive to zero-query settings.
\input{ds-correctness}

\heading{Privacy. }
The privacy of $\setprim_{\mathrm{ds}}$ follows directly from our general result, Theorem~\ref{thm:priv-ss}.
\begin{theorem}[\rm{SS-privacy of DS PRF}]\label{thm:ds-ss}
Fix $k,m,n>0$, a set~$\elts$, and let $\setprim_{\mathrm{ds}}= (\Rep, \Qry)$ be the set-membership data structure described in Figure~\ref{fig:bf-and-ds}.  For any set~$S \in 2^\elts$, let $\leak(S)=|S|$.  There exists a simulator~$\Sim$ and adversary~$B$ such that, for any adversary~$A$
\[
\AdvPrivSS{\setprim_{\mathrm{ds}},\leak}{A,\Sim} \leq  \AdvPRF{F}{B}\;.
\]
%(Both~$B$ and~$\Sim$ are explicitly constructed in the proof of this theorem.)
\end{theorem}
\begin{proof}
It suffices to give the decomposition of $\Rep$ in to $\Rep_1$ and $\Rep_2$ as required by Theorem~\ref{thm:priv-ss}.
On input a set of distinct elements $S=\{x_1,x_2,\ldots,x_n\}$, algorithm~$\Rep_1$ returns $s_1=\langle 1,x_1 \rangle, s_2=\langle 2,x_1 \rangle, \ldots, s_k\langle k,x_1\rangle, s_{k+1}=\langle 1,x_2\rangle, \ldots,\linebreak s_{2k}=\langle k,x_2\rangle, \ldots ,s_{kn+1}=\langle 1,x_n\rangle, \ldots, s_{kn}=\langle k,x_n \rangle$.  On input $n,F_K(s_1),\ldots,\linebreak F_K(s_{kn})$, algorithm $\Rep_2$ simply initializes an all-zero array~$M$ of~$m$-bits, and sets the positions $M[F_K(s_i)]$ to 1.  It returns~$M$ as $\pubaux$. \hfill\qed
\end{proof}

%%%%%%%%%%%%%%%%% LINEAR-PRF %%%%%%%%%%%%%%%%%%%%
\begin{figure}[tp]
\centering
\hfpagess{.5}{.5}
{
\algorithmv{$\Rep(S)$}\\
$M \gets 0^m$; $K \getsr \calK$\\
for $x \in S$\\
\jnote{Fix}\tsnote{Fix what?}\\
\nudge for $j \in \{1,2,\ldots,k\}$\\
\nudge\nudge $h_j \gets F_{K}(\langle 1,x \rangle) + j\cdot F_{K}(\langle 2,x \rangle) \bmod m$\\
\nudge\nudge $M[h_j] \gets 1$\\
$\pubaux \gets M$, $\privaux \gets \langle K \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$M\gets\pubaux$, $K \gets \privaux$\\
\jnote{Fix}\tsnote{Fix what?} \\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets F_{K}(\langle 1,x \rangle) + j\cdot F_{K}(\langle 2,x\rangle) \bmod m$\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1
}
{
\algorithmv{$\Rep(S)$}\\
$\pubaux \gets \emptystring$\\
$K_1,K_2, \getsr \calK$\\
for each $x \in S$\\
\nudge $M_x \gets 0^m$\\
\nudge $B_x \gets \mathsf{bigram}(x)$\\
\nudge for each $b \in B_x$\\
\nudge\nudge for $j \in [k]$\\
\nudge\nudge\nudge $i \gets F_{K}(\langle 1, b\rangle)+j\cdot
F_{K}(\langle 2, b  \rangle) \bmod m$\\
\nudge\nudge\nudge $M_x[i] \gets 1$\\
\nudge $\pubaux \gets \langle\pubaux, M_x \rangle$\\
$\privaux \gets \langle K_1,K_2 \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$B \gets \mathsf{bigram}(x)$\\
$\overline{M} \gets 0^m$\\
for each $b \in B$\\
\nudge for $j \in [k]$\\
\nudge\nudge $i \gets F_{K}(\langle 1,b \rangle)+j\cdot F_{K}(\langle 2,b \rangle) \bmod m$\\
\nudge\nudge $\overline{M}[i] \gets 1$\\
if $\overline{M} \in \pubaux$ then Return 1 \\
Return 0
}
\caption{{\bf Left:} $\Rep$ and $\Qry$ algorithms for the linear-PRF
  data structure $\setprim_{\mathrm{lin}}$ {\bf Right:}
The scheme from Niedermeyer et al.\cite{xxx}, $\setprim_{\bilin}$. The $\mathsf{bigram}$
algorithm takes a string~$x=x_1,x_2,\ldots,x_\ell$ as input, and
returns the set of bigrams for the left- and right-padded string
$\sqcup x \sqcup$, i.e. $\{\sqcup x_1, x_1x_2,\ldots,x_{\ell-1}x_\ell,x_\ell\sqcup\}$.
}
\label{fig:lin-and-bi-lin}
\label{fig:niedermeyer}
\end{figure}

\subsection{Linear PRF-hash Bloom Filter }
%\heading{Linear-PRF construction. }
The domain-separated PRF-hash requires~$k$ PRF evaluations per set-membership query, and this may be unattractive in certain use cases.  Kirsch and Mitzenmacher~\cite{kirsch2006less} have shown that two PRF evaluations suffices, defining the hash functions to be $h_i(x)=F_K(\langle 1,x \rangle) + j\cdot F_K(\langle 2,x \rangle) \bmod m$, where~$m$ is the bitlength of the Bloom filter.\footnote{Actually, they define $h_i(x)=F_{K_1}(x)  + j\cdot F_{K_2}(x) \bmod m$, using distinct keys instead of domain-separation, but the constructions are equivalent with respect to security.} We call this the \emph{linear PRF-hash} Bloom Filter.  In this section, we show that this scheme is both correct and private with respect to our SS notion.

Looking ahead to the next section, we will analyze a closely related construction, which was attacked by Niedermeyer et al.~\cite{niedermeyer2014cryptanalysis} in the context of private medical-record linkage.  It uses the same linear-PRF hash functions, but has a $\Rep$ algorithm that that does not ensure that the hash functions are called on \emph{distinct} values.  We show that this is \emph{not} private with respect to our SS notion for a leakage function that leaks nearly the entire set~$S$. \todo{It would be better, perhaps, to show that it is not OW secure, since this implies it isn't SS secure, and OW security is what Niedermeyer et al. violate.}

Taken together, the results of this section and the next highlight how sensitive the privacy of linear PRF-hash constructions is to usage.

\heading{Correctness. }
For now, let $n,k,m \geq 0$.  Then $\setprim_{\mathrm{lin}}= (\Rep, \Qry)$ is defined as in Figure~\ref{fig:lin-and-bi-lin} (left side).  The following result shows, informally, that if~$F$ is a good PRF, then $\setprim_{\mathrm{lin}}$ is correct against adaptive error-finding adversaries.

\begin{theorem}[{\rm Correctness of linear-PRF}]\label{thm1}\label{thm:lin-correctness}
Fix $k,m,n,r>0$, a set~$\elts$, and let $\setprim_{\mathrm{lin}}= (\Rep, \Qry)$ be the set-membership data structure (over~$\elts$) described in Figure~\ref{fig:lin-and-bi-lin}.  For any adversary~$A$ that makes a total of~$q$ queries to its oracle and has time-complexity~$O(t)$, there exist prf-adversaries~$B_1,B_2$ (explicitly constructed in the proof of this theorem) such that
\[
\AdvCorrect{\setprim_{\mathrm{lin}},r}{A} \leq  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}  +{\dbinom{q}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B_1$ and $B_2$, each ask~$q$ oracle queries and have time complexity~$O(t+qm)$.
\end{theorem}
Again, Kirsch and Mitzenmacher~\cite{kirsch2006less} show
that the ($r=1$, zero-query) error-probability for~$\setprim_\mathrm{lin}$ is
always within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic
rate of convergence is faster than~$1/n$.   
the multiplicative term $\binom{q}{r}$ arises when switching from the adaptive to zero-query settings.
\input{lin-correctness}

\heading{Privacy. }
The SS-privacy of the linear-PRF hash construction follows directly from~Theorem~\ref{thm:priv-ss}.
\begin{theorem}[{\rm SS-privacy of linear-PRF}]\label{thm:lin-privacy}
Fix $k,m,n,r>0$, a set $\elts$, and let $\setprim_{\mathrm{lin}}= (\Rep, \Qry)$ be the set-membership data structure (over~$\elts$) described in Figure~\ref{fig:lin-and-bi-lin}. For any set~$S \in 2^{\elts}$, let $\leak(S)=|S|$.  There exists a simulator~$\Sim$ and an adversary~$B$ such that, for any adversary~$A$
\[
\AdvPrivSS{\setprim_{\mathrm{ds}},\leak}{A,\Sim} \leq  \AdvPRF{F}{B}\;.
\]
%(Both~$B$ and~$\Sim$ are explicitly constructed in the proof of this theorem.)
\end{theorem}
\begin{proof}
It suffices to give the decomposition of $\Rep$ in to $\Rep_1$ and $\Rep_2$ as required by Theorem~\ref{thm:priv-ss}.
On input a set of distinct elements
$S=\{x_1,x_2,\ldots,x_n\}$, algorithm~$\Rep_1$ returns $s_1=\langle 1,x_1 \rangle, s_2=\langle 2,x_1 \rangle, s_3 = \langle 1,x_2\rangle, s_4=\langle 2,x_2 \rangle,\ldots,\linebreak s_{2n-1}=\langle 1,x_n \rangle, s_{2n}=\langle 2,x_n \rangle$.

On input $n,F_K(s_1),F_K(s_2),\ldots,F_K(s_{2n-1}) , F_K(s_{2n})$, algorithm $\Rep_2$ initializes an all-zero array~$M$ of~$m$-bits, and sets the positions $M[F_K(s_i)+j\cdot F_K(s_{i+1})\bmod m]$ to 1, for each $j \in [k]$.  It returns~$M$ as $\pubaux$. \hfill\qed
\end{proof}

%%%%%%%%%%%%%%%%% BIGRAM LINEAR-PRF (AKA "NEIDERMAYER")  %%%%%%%%%%%%%%%%%%%%
\subsection{Bigram Linear-PRF Construction}
%\heading{Bigram Linear-PRF construction (``Neidermayer''). }
%\heading{Correctness.}
%\todo{Correctness theorem for this scheme.  Should follow pretty closely the one for ``plain'' linear-PRF.}\tsnote{Actually not sure this is that interesting, but for parallel structure one would want to give it.}
\heading{Privacy. }
Consider the set-membership data structure given in Figure~\ref{fig:lin-and-bi-lin} (right side), which we call the \emph{bigram linear-PRF} construction.  This construction is of practical interest, as it has been suggested as a way to address the privacy-preserving record-linkage problem, in particular for linking common records in medical databases~\cite{niedermeyer2014cryptanalysis,schnell2011novel}.
We only address the privacy of this construction, as the correctness follows immediately from the correctness of the Linear-PRF contruction.  The privacy, however, does not.  Here we exhibit an efficient SS-privacy attack even when simulator is provided nearly the entire set.  Consider the following ``all-but-last'' leakage function.  On input a set~$S=\{x_1,x_2,\ldots,x_n\}$, it assigns $x'_n \gets x_n[1:|x_n|-1]$ and returns $\mathsf{leak_{abl}}(S)=\{x_1,x_2,\ldots,x'_n\}$.  \todo{OW privacy proof? Or just some text that says that Niedermeyer et al show something stronger~---OW-privacy violation---~albeit with $n=10000, m=10000\alpha$ for an $\alpha$ that I need to look up.}

\begin{figure}[tp]
\centering
\fpage{.65}
{
\adversaryv{$A$}\\[1ex]
On input $\emptystring$:\\
\nudge $x_1 \gets \mathrm{cab}$\\
\nudge $x_2 \gets \mathrm{bab}$\\
\nudge $z \getsr \bits$\\
\nudge if $z=1$ then $x_3 \gets \mathrm{caba}$\\
\nudge else $x_3 \gets \mathrm{cabc}$\\
\nudge Return $S=\{x_1,x_2,x_3\}$\\
On input $\pub \neq \emptystring$:\\
\nudge $M_1,M_2,M_3 \gets \pub$\\
\nudge $M_{1,2} \gets M_1 \vee M_2$\\
\nudge if $z=1 \wedge \left(3k < \mathsf{hamming}(M_{1,2} \wedge M_3) = 4k\right)$ then \\
\nudge\nudge Return 1 \\
\nudge if $z=0 \wedge \left(\mathsf{hamming}(M_{1,2} \wedge M_3) \leq 3k\right)$ then \\
\nudge\nudge Return 1 \\
\nudge Return 0
}
\caption{Adversary for Theorem~\ref{thm:bi-lin-ss}. The function $\mathsf{hamming}(M)$ returns the hamming weight of~$M$.}
\label{fig:adv-bi-lin-ss}
\end{figure}

\begin{theorem}[{\rm SS-privacy of bigram linear-PRF}]\label{thm:bi-lin-ss}
Let $\setprim_{\bilin}= (\Rep, \Qry)$ be the set-membership data structure described in Figure~\ref{fig:neidermayer}, but instantiated with $F_{K}$ replaced by a random functions~$\rho$.  Let $\leak_{\mathsf{abl}}$ be the leakage function just described.   Let~$\elts=\{\mathrm{a,b,c}\}^*$. Let $n=3$,  $m=3\alpha $ for integer $\alpha >> 1$, and let, $k = \lceil \ln(2) \alpha \rceil$.\tsnote{This value of~$k$ is from the classical analysis for minimizing the FP rate for a Bloom filter, and is set this way for compatibility with classical results.  It isn't used in any essential way in the proof sketch.}  Let~$A$ be the adversary in Figure~\ref{fig:adv-bi-lin-ss}.  Then for any simulator $\Sim$, the advantage $\AdvPrivSS{\setprim_{\bilin},\leak_{\mathsf{abl}}}{A,\Sim}$ can be made arbitrarily close to $1/2$ by choice of~$\alpha$.
\end{theorem}
\begin{proof}\tsnote{still a bit sketchy but believable}
On input the set~$S$ produced by~$A$, the $\Rep$ algorithm turns $x_1=$'cab', $x_2=$'bab' and $x_3$ into their respective bigram sets: $B_{x_1}=\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{b}\sqcup\}$, $B_{x_2}=\{\sqcup\mathrm{b},\mathrm{ba},\mathrm{ab}, \mathrm{b}\sqcup \}$; the set $B_{x_3}$ will be either $\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{ba}, \mathrm{a}\sqcup\}$, or $\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{bc}, \mathrm{c}\sqcup\}$, each with probability 1/2.  It then produces $\alpha$-bit arrays $M_1, M_2, M_3$ by hashing the corresponding bigram sets using the hash functions $h_j(s)=\rho_1(s)+ j\rho_2(s) \bmod m$.  Note that for sufficiently large~$m$, each of $M_1,M_2$ will contain exactly $4k$ positions set to 1; this is with overwhelming probability, up to collisions in the hash functions for different bigrams within a given bigram set.  Likewise, $M_{1,2}=M_1 \vee M_2$ will have exactly $6k$ positions set to 1, because $B_{x_1}$ and $B_{x_2}$ share $\mathrm{ab}$ and $\mathrm{a}\sqcup$.

In the case that $B_{x_3}=\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{ba}, \mathrm{a}\sqcup\}$, only one of its elements ($\mathrm{a}\sqcup$) is not in either of $B_{x_1},B_{x_2}$.  Hence, the hamming weight of $M_{1,2}\wedge M_3$ will be $4k$.  On the other hand, when $B_{x_3} = \{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{bc}, \mathrm{c}\sqcup\}$, two of its elements ($\mathrm{bc}, \mathrm{c}\sqcup$) are not covered by $B_{x_1},B_{x_2}$.  In this case, the hamming weight of $M_{1,2} \wedge M_3$ will be $3k$.  Thus $\ExpPrivSSreal{\setprim_{\bilin}}{A}=1$ with overwhelming probability.

On the other hand, the simulator~$\Sim$ is given 'cab','bab','cab' as
input.  It can perfectly simulate $M_1,M_2$ by lazily sampling random
function~$\rho$ itself.  However, it has no information about the
bit~$z$, so at best it can cause~$A$ to output 1 with
probabilility~$1/2$.  That is $\ExpPrivSSsim{\setprim_{\bilin}}{A,\Sim} \leq 1/2$. \hfill\qed
\end{proof}

%%%%%%%%%%%%%%%%% Bigram DOMAIN-SEPARATED PRF %%%%%%%%%%%%%%%%%%%%
\subsection{Bigram Domain-Separated PRF-Hash}
We note that Niedermeyer et al. in~\cite{niedermeyer2014cryptanalysis} suggest replacing the linear-PRF hash functions in the previous construction with independently keyed PRFs as a countermeasure to their attack.  In terms of SS-privacy, however, this change offers no help.  The following theorem is proved via a nearly identical argument as Theorem~\ref{thm:bi-lin-ss}.

\begin{theorem}[{\rm SS-privacy of bigram DS PRF-hash}]\label{thm:bi-ds-ss}
Let $\setprim_{\bilin}= (\Rep, \Qry)$ be the set-membership data structure described in Figure~\ref{fig:lin-and-bi-lin}, but instantiated with $F_{K}$ replaced by a random functions~$\rho$.  Let $\leak$ be the leakage function described just prior to Theorem~\ref{thm:bi-lin-ss}.   Let~$\elts=\{\mathrm{a,b,c}\}^*$. Let $n=3$,  $m=3\alpha $ for integer $\alpha >> 1$, and let, $k = \lceil \ln(2) \alpha \rceil$.\tsnote{This value of~$k$ is from the classical analysis for minimizing the FP rate for a Bloom filter, and is set this way for compatibility with classical results.  It isn't used in any essential way in the proof sketch.}  Let~$A$ be the adversary in Figure~\ref{fig:adv-bi-lin-ss}.  Then for any simulator $\Sim$,
the advantage $\AdvPrivSS{\setprim_{\mathrm{ds}},\leak}{A,\Sim}$ can be made arbitrarily close to $1/2$ by choice of~$\alpha$.
\end{theorem}

\todo{OW privacy proof? Attack? [JON]}

%%%%%%%%%%%%%%%%% GARBLED BF  %%%%%%%%%%%%%%%%%%%%
\subsection{Garbled Bloom Filter}
\begin{figure}[tp]
\centering
\fpage{.8}{
\hpagess{.55}{.45}
{
\algorithmv{$\Rep^H(S)$}\\
$M \gets (\bot)^m$\\
%$K_1,K_2,\ldots, K_k \getsr \calK$\\
for $x \in S$\\
\nudge $\finalshare \gets 0^\lambda$; $\emptyslot \gets -1$\\
\nudge for $j \in \{1,2,\ldots, k\}$\\
\nudge \nudge $h_j \gets H(j,x)$\\
\nudge \nudge if $M[h_j] \neq \bot$ then\\
\nudge \nudge \nudge $\finalshare \gets \finalshare \xor M[h_j]$\\
\nudge \nudge else if $\emptyslot = -1$ then\\
\nudge \nudge \nudge $\emptyslot \gets h_j$\\
\nudge \nudge else\\
\nudge \nudge \nudge $M[h_j] \getsr \bits^\lambda$\\
\nudge \nudge \nudge $\finalshare \gets \finalshare \xor M[h_j]$\\
\nudge $M[\emptyslot] \gets \finalshare \xor x$\\
for $i \in \{1,2,\ldots, m\}$\\
\nudge if $M[i] = \undefined$ then\\
\nudge \nudge  $M[i] \getsr \bits^\lambda$\\
$\pubaux \gets \emptystring$; $\privaux \gets \langle M\rangle$ \mbox{ (if private~$M$)}\\
$\pubaux \gets \langle M \rangle$; $\privaux \gets \emptystring$ \mbox{(if public~$M$)}\\
}
{
\algorithmv{$\Qry^H(\pubaux,\privaux,\qry_x)$}\\
if $\privaux \neq \emptystring$ then $M \gets \privaux$\\
else $M\gets \pubaux$\\
$x' \gets 0^\lambda$\\
for $j \in \{1,2,\dots,k\}$\\
\nudge $h_j \gets H(j,x)$\\
\nudge $x' \gets x' \xor M[h_j]$\\
if $x' =x $ then Return 1\\
Return 0
}
}
\caption{ $\Rep$ and $\Qry$ algorithms for garbled Bloom
  filter $\setprim_{\mathrm{garbled}}$ in the ROM.hash function $H\colon \calK \times
 \elts \to [m]$. The query set is $\calQ=\{\qry_x\colon
  \univ \to \bits\}$ where $\qry_x(S)=1 \Leftrightarrow x \in S$.}
\label{fig:gbf}
\end{figure}

\begin{figure}[tp]
\centering
\hfpagess{.45}{.45}
{
\adversaryv{$A_{\mathrm{ss}}^{\TestOracle,H}(\pubaux)$}\\[1ex]
$M \gets \pubaux$\\
for $\ell=1$ to $\binom{m}{k}$\\
\nudge $i_1,i_2,\ldots,i_k \gets \mathcal{I}_\ell$\\
\nudge $x \gets M[i_1] \xor M[i_2] \xor \cdots \xor M[i_k]$\\
\nudge if  $x \not\in S \wedge \left(\bigwedge_{j=1}^k H(j,x)=i_j \right)$ then \\
\nudge\nudge $a\gets\TestOracle(x)$\\
\nudge\nudge Return $\bot$\\
}
{
\adversaryv{$A_{\mathrm{ow}}^{H}(\pubaux)$}\\[1ex]
$M \gets \pubaux$\\
for $x \in X_1,X_2,\ldots,X_{q-1}$\\
\nudge for $j=1$ to $k$ \\
\nudge \nudge \nudge $i_j \gets H(j,x)$\\
\nudge $x' \gets M[i_1] \xor M[i_2] \xor \cdots \xor M[i_k]$\\
\nudge if $x' = x$ then Return $x$\\
Return~$X_{q}$
}
%{
%\adversaryv{$A^{\TestOracle}(\pubaux)$}\\[1ex]
%$M \gets \pubaux$\\
%for $1 \leq i_1 < i_2 < \cdots < i_k \leq m$\\
%\nudge $x \gets M[i_1] \xor M[i_2] \xor \cdots \xor M[i_k]$\\
%\nudge if $\TestOracle(x)=1$ then Return $\bot$
%}
%\caption{Adversary for Theorem~\ref{thm:xxx}.  \textbf{Left:} public hash case, \textbf{Right:} private hash case.}
\caption{{\bf Left:} Correctness attack for garbled Bloom filters with public~$M$.  The sequence $\mathcal{I}_1,\mathcal{I}_2,\ldots$ is some implicit enumeration of~$k$-tuples of indices over $[m]$.
{\bf Right:} One-wayness attack for garbled Bloom filters with public~$M$.  The sequence $X_1,X_2,\ldots,X_q,$ are the~$q$ most probable points in the support of~$\distr{}{}$ in decreasing order.  }
\label{fig:adv-gbf}
\end{figure}
Dong, Chen and Wen~\cite{dong2013private} introduce a set-membership data
structure called a \emph{garbled Bloom filter} (GBF).   Modulo some implementation
details, a GBF is like a classical Bloom filter, but
rather than storing a bit 1 at the indices determined by $H_j(x)$, the
GBF stores xor-shares of~$x$.  Thus, if $x \in S$, then hashing~$x$
and xoring together the strings stored at the indicated positions
recovers~$x$. We give the pseudocode for the GBF $\Rep$ and $\Qry$
algorithms in Figure~\ref{fig:gbf}.

In~\cite{dong2013private} the GBF is a central piece of a private set-intersection
scheme, in which the client computes a classical BF over its
set~$S_1$, the server computes a GBF over its set~$S_2$, and then an
oblivious transfer protocol is carried out, at the end of which the
client obtains a GBF for $S_1 \cap S_2$ (the server learns nothing).
The authors note that the GBF is a new set-membership data strucure
that may be of general interest, so we explore its correctness and
privacy with respect to use case that may or may not make the
representation~$M$ public. The pseudocode in Figure~\ref{fig:gbf}
covers both possibilities.

\heading{Correctness.}
We note that~\cite{dong2013private} provides error-rate analysis, i.e., the
probability of finding a single false-positive in a non-adaptive
attack.  When $\lambda$ is the bit-length of the xor-shares, they
argue that the error-rate is at most $2^{-\lambda}$, roughly because
the xor-shares stored in~$M$ are random strings.  (Note that
  this is independent of~$k,m,n$; although $k>1$ is required for the
  xor shares.)  However, this analysis implicilty assumes that the potential false-positive string~$x$ is fixed
\emph{before}~$M$ is determined.  We will see that this carries over
\tsnote{I think?} to the case in which~$M$ is private (i.e.,
$\pubaux=\emptystring,\privaux=\langle M \rangle$), but when~$M$ is
public things are different.

Consider the adversary described in the left half of Figure~\ref{fig:adv-gbf}.  In one
round of the for-loop, it pics~$k$ positions $i_1,i_2,\ldots,i_k$, and
xors the strings at these positions to get a candidate~$x$.   If $x
\in S$, which the adversary knows in the correctness experiment, then
it moves on to a new set of positions.  If $x \not\in S$ and
$h_{1}(x)=i_1 \wedge h_2(x)=i_2 \wedge \cdots \wedge h_k(x) = i_k$,
then it \emph{must be} that $x$ is a false-positive.   Indeed, note
that the probability of this (in the ROM) is precisely the
false-positive rate of a classical BF using the same $(k,m,n)$ as the GBF.

Now, this adversary will find a false positive, if there is one, but in the worst case it will consider
$\binom{m}{k} \approx m^k$ $k$-tuples of positions.  If both~$m$
and~$k$ are large this is infeasible.  In~\cite{dong2013private} they suggest to
set~$k=\lambda$, and test with $\lambda=80$.  Making 80 hash calls per
set-membership query would be prohibitive in many applications, and a
$2^{-80}$ false-postive rate would be typically considered overkill in
practice.\footnote{Note
that, as described in~\cite{xxx}, $\lambda$ must be at least the
bitlength of the longest element $x \in S$. \tsnote{This could be
  handled with a slight change to the set membership test, by checking
if the $\lambda$-bit shares xor to, say, the $\lambda$-bit truncation
of~$x$.}}
Let us consider a practically interesting set of parameters. Say you have a 1000-position
filter, $m=1000$, and a 100-element set, $n=100$,  and $k=4$ hash
functions (a la Squid\cite{fan2000summary}).  This results in a false-positive rate of
approximately 1\% in a classical BF.  There are roughly $2^{35}$
subsets of four indices to consider, which is reasonable on a
commodity laptop.  Moreover, with $n=100$ the expected
number of trials is roughly $2^{28}$.
Note that the attack is easily extended to find $r>1$ false positives.

We admit that attack is not particularly sophisticated, yet it might
be significantly more efficient than brute force guessing.  When $k,m,n$ are as in the example just given, and $\elts=\bits^{128}$, 
this attack will be considerably faster even if the attacker knows that the distribution deviates significantly from uniform.   Note that in the ROM, the order in which $k$-tuples are considered is largely irrelevant, especially when $2^\lambda \gg \binom{m}{k}$.
%Notice,too, that you \emph{cannot} mount it against a standard Bloom filter.

Finally, we observe that a similar version of the correctness attack
works outside of the ROM, say, when we hash via $H(j,x)=F_K(\langle
j,x \rangle)$ and assume~$F_K$ is a secure PRF.  Here,
$\privaux=\langle K \rangle$.  Now,
however, the work switches from offline hash evaluations (i.e., random
oracle calls) to online set-membership (i.e., $\TestOracle$) queries.

Switching now to the case of a \emph{private}-$M$ GBF, i.e., $\pubaux=\emptystring, \privaux=\langle M \rangle$, we have the
following correctness bound.

\begin{theorem}[{\rm Correctness of private-$M$ GBF in the ROM}]\fixme{FINISH! (Animesh?)}\label{thm:gbf-correctness}
Fix $k,m,n,r,\lambda>0$, and let $\setprim_{\mathrm{gbf}}=(\Rep, \Qry)$ be the private-$M$ garbled Bloom filter. For any adversary~$A$ that makes a total of~$q_T$ queries to the $\Test$ oracle, and $q_R$ queries to the RO, and has time-complexity~$O(t)$,
\[
\AdvCorrect{\setprim_{\mathrm{gbf}},r}{A} \leq  {\dbinom{q_T}{r}} \left( \dfrac{1}{2^{\lambda}} \right)^r\,.
\]
\end{theorem}

\heading{Privacy. }
We observe immediately that when~$\pubaux=\langle M \rangle$ the GBF cannot be SS-private (in the ROM) because $\privaux=\emptystring$.  In fact, the correctness attack given above suggests a one-wayness attack in the ROM, too; see the right half of Figure~\ref{fig:adv-gbf}.   The naive attack would simply output the most probable element in the support of the distribution $\distr{}{}$ used in the one-wayness experiment to sample~$S$.  But when~$M$ and the hash functions are available to the attacker, it can carry out set-membership tests for itself.  Thus the adversary on the right of Figure~\ref{fig:adv-gbf} tries the~$q-1$ most probable points, in order, halting with its output as soon as the set-membership test is successful.  If none of these points are yield a successful test, then the next most probable point is output.  It is not hard to conclude the following theorem.

\begin{theorem}[{\rm OW-privacy of public-$M$ GBF}]
Fix $k,m,n,r,\lambda>0$, and let $\setprim_{\mathrm{gbf}}=(\Rep, \Qry)$ be the public-$M$ garbled Bloom filter, in the random-oracle model for~$H$.  Then \fixme{This probability might not be correct... verify when less tired.}
\[
\AdvPrivOW{\setprim_{\mathrm{gbf}},\distr{}{}}{A_{\mathrm{ow}}} \geq 1 - \left(1-\frac{n}{2^{\mu}} \right)^{q} \geq 1 - e^{-nq/2^\mu}
\]
 where $\mu = H_\infty(\distr{}{})$ is the pointwise min-entropy of~$\distr{}{}$, and $A_{\mathrm{ow}}$ makes $kq$ random-oracle queries.
\end{theorem}
\noindent
We conjecture that this is the optimal attack on the GBF in this setting.

If instantiated with hash functions $H_j(x) = F_K(\langle j,x \rangle)$ and $\privaux = \langle K \rangle$, the SS-privacy of the GBF follows directly from~Theorem~\ref{thm:priv-ss}.
\begin{theorem}[{\rm SS-privacy of GBF with secret hash functions}]
\label{thm:gbf-prf-privacy}
Fix $k,m,n,r,\lambda>0$ and a function family~$F\colon \calK \times \bits^* \to [m]$.  Let $\setprim_{\mathrm{gbf\mbox{-}prf}}= (\Rep, \Qry)$ be the garbled Bloom filter with hash functions $H_j(x)=F_K(\langle j,x \rangle)$ and $\privaux=\langle K \rangle$.  For any set~$S$, let $\leak(S)=|S|$.  There exists a simulator~$\Sim$ and an adversary~$B$ such that, for any adversary~$A$
\[
\AdvPrivSS{\setprim_{\mathrm{gbf\mbox{-}prf}},\leak}{A,\Sim} \leq  \AdvPRF{F}{B}\;.
\]
%(Both~$B$ and~$\Sim$ are explicitly constructed in the proof of this theorem.)
\end{theorem}
\begin{proof}
It suffices to give the decomposition of $\Rep$ in to $\Rep_1$ and $\Rep_2$ as required by Theorem~\ref{thm:priv-ss}.
On input a set of distinct elements $S=\{x_1,x_2,\ldots,x_n\}$, algorithm~$\Rep_1$ returns $s_1=\langle 1,x_1 \rangle, s_2=\langle 2,x_1 \rangle, \ldots, s_k\langle k,x_1\rangle, s_{k+1}=\langle 1,x_2\rangle, \ldots,\linebreak s_{2k}=\langle k,x_2\rangle, \ldots ,s_{kn+1}=\langle 1,x_n\rangle, \ldots, s_{kn}=\langle k,x_n \rangle$.  On input $n,F_K(s_1),\ldots,\linebreak F_K(s_{kn})$, algorithm $\Rep_2$ carries out the pseudocode of~$\Rep$ in Figure~\ref{fig:gbf}, making use of $F_K(s_1),\ldots,F_K(s_{kn})$ in the obvious way.  It returns~$M$ as $\pubaux$.\hfill\qed
\end{proof}

\ignore{
\begin{verbatim}
%----- LEFT HERE FOR REMINDERS, AND FOR MENTIONING OPEN PROBLEMS IN INTRO-------------------------------%
\end{verbatim}

\todo{Below are two lists of things that we should show.}

\heading{Correctness. }  Here we give security proofs and attacks with respect to the correctness notions.
\begin{itemize}
\item Show Bloom Filter is correct in the ROM. \jnote{Should follow along the lines
of Theorem~2, but taking into account adversary's oracle queries.} [DONE]
\item correctness security/attacks on the bigram/PRF-based double-hashing construction ($h_j(x) = F_{K1}(x) + j\cdot F_{K2}(x) \bmod m$ where the~$x$ are bigrams of names, etc.)
\item correctness-security/attacks on the bigram/domain-separated PRF construction ($h_j(x)=F_K(\langle j,x \rangle) \bmod m$, or just assume range of~$F$ is $[m]$.)
\item correctness-security/attacks Dong, Chen, Wen ``garbled Bloom filter'' construction

\end{itemize}

\heading{Privacy. }
\begin{itemize}
\item Prove that the basic BF, in the ROM, is private.  \tsnote{Not SS, should be OW}
\item Show that the bigram DS-PRF construction is/isn't SS [DONE]
\item Show that the basic BF DS-PRF construction is SS [DONE]
\item Prove privacy bounds for the constructions from NY \tsnote{Follows from Jon's meta-theorem}

\item Cast Neidermayer et al.\ attacks on the bigram/PRF-based double-hashing construction (see their paper) into our formalism.  \tsnote{Note that their attack would allow one to recover \emph{every} name, not just one.}

\item Prove a privacy bound for the bigram/PRF-based double-hashing construction that Neidermayer et al.\ attacks. I expect that the provable security bound is considerably worse than the min-entropy because of the way the representation is made.  (Essentially, the Hamming weight of the representation is a good estimate of the length of the longest surname in filter.)
\jnote{Maybe tailor leakage function appropriately, i.e., so that it only leaks the number of bigrams?}

\item Attack correctness of Dong, Chen, Wen ``garbled Bloom Filter'' construction when the hash functions are public.\tsnote{I give the attack in the related work section.  Copy it here and flesh it out.}  Prove priv-security privacy of construction when hash functions are secret.  \tsnote{Makes the point that the application setting really matters.}

\item Prove privacy bounds of other suggestions, such as the record-level BF from Durham et al.?  This would be a nice pairing with the Neidermayer et al.\ results \tsnote{No time.}

\end{itemize}
}