### 2016-10-12 EC17-submitted/Theorem3_priv.tex

\begin{theorem}\label{thm3}
A classic Bloom filter is private in random oracle model(ROM), i.e. an adaptive adversary $A$ has low probability of winning the $\ExpPrivPubHashBB{B}{\distr{\univ}{n},A}$ game in ROM. For $q$ queries and min-entropy ($\mu$) of $\distr{\univ}{n}$,

\begin{equation}
\AdvPrivPubHashBB{B}{A} \leq  \frac{q}{2^\mu} + \frac{n}{\abs{\univ}}
\end{equation}

\end{theorem}

\begin{proof}[\Cref{thm3}]
As shown in \Cref{fig:GameT3}, $\game{0}(A)$ and $\game{1}(A)$ exactly simulate $\ExpPrivPubHashBB{B}{\distr{\univ}{n}, A}$ in ROM, except in $\game{1}(A)$, the random oracle always returns $k$ uniformly random values for any $x \in \univ$.
\begin{equation}
\AdvPrivPubHashBB{B}{A} = \Prob{\game{0}(A)=1}\label{eq:0T3}
\end{equation}
%\caption{Game playing argument}\label{fig:Game}
\begin{figure}
\centering
\fpage{.6}{
\hpagess{.4}{.45}
{
$\game{1}(A)$\\
\fbox{$\game{0}(A)$\\
}
$\\ \calS \getsr \distr{\univ}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$x \getsr A^{\MemOracle,\QryOracle,\HashOracle}(M)$\\
if $x\in \calS$ then Ret 1\\
Ret 0
}
%
{
\oracle{$\MemOracle(x)$}\\
Ret $[x \in \calS]$\\

\medskip
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\

\medskip
\oracle{$\HashOracle(x)$}\\
$\rho_i \getsr Func(\univ,[m])$, for $i\in[1,k]$\\
$z_{i} \gets \rho_i(x)$, for $i\in[1,k]$\\
if $T[x] \neq $ undefined\\
\nudge Bad $\gets$ true\\
\fbox{
$z_i \gets T[x]$, for $i\in[1,k]$
}
$\\T[x] \gets z_i$, for $i\in[1,k]$\\
Ret $z_i$, for $i\in[1,k]$\\
}
}
\caption{\Cref{thm3}: Game playing argument}\label{fig:GameT3}
\end{figure}

\noindent
It is evident from $\game{0}(A)$ and $\game{1}(A)$ that 
\begin{equation}
\Prob{\game{0}(A) : \mathrm{Bad}=\mathrm{true}} = \Prob{\game{1}(A) : \mathrm{Bad}=\mathrm{true}}\label{eq:1T3}
\end{equation}

\noindent
Using fundamental lemma of game playing argument and \Cref{eq:1T3}
\begin{eqnarray}
\nonumber \Prob{\game{0}(A)=1} & \leq & \Prob{\game{0}(A)=1 | \mathrm{Bad} \neq \mathrm{true}} + \Prob{\game{0}(A) : \mathrm{Bad}=\mathrm{true}}\\
&=& \Prob{\game{1}(A)=1} + \Prob{\game{1}(A) : \mathrm{Bad}=\mathrm{true}}\label{eq:2T3}
\end{eqnarray}

\noindent
Let the random oracle be modified to output $k$ uniformly random values without setting any Bad flag. Let's call it $\game{2}(A)$ which is shown in \Cref{fig:Game2T3}. It turns out that the  $\QryOracle$ is no more deterministic due to coins of the random oracle. Hence, the results of the $q$ queries by $A$ are mutually independent and random. So, neglecting the output of $A$, the game samples one of the elements queried by $A$, and returns 1/0.

\begin{equation}
\Prob{\game{1}(A)=1} = \Prob{\game{2}(A)=1} = \frac{q}{2^\mu}\label{eq:3T3}
\end{equation}

\begin{figure}
\centering
\fpage{.6}{
\hpagess{.5}{.5}
{
\underline{$\game{2}(A)$}\\
$\calZ \gets \emptyset$\\
$\calS \getsr \distr{\univ}{n}$\\
$(M,\tau) \getsr \Rep^{\HashOracle}(\calS)$\\
$x' \gets A^{\MemOracle,\QryOracle,\HashOracle}(M)$\\
$x \getsr \calZ$\\
if $x\in \calS $ then Ret 1\\
Ret 0
}
%
{
\oracle{$\MemOracle(x)$}\\
Ret $[x \in \calS]$\\

\medskip
\oracle{$\QryOracle(x)$}\\
Ret $\Qry^{\HashOracle}(M,\tau,x)$\\

\medskip
\oracle{$\HashOracle(x)$}\\
$\calZ \gets \calZ \cup x$\\ 
$\rho_i \getsr Func(\univ,[m])$, for $i\in[1,k]$\\
$z_i \gets \rho_i(x)$, for $i\in[1,k]$\\
Ret $z_i$, for $i\in[1,k]$\\
}
}
\caption{\Cref{thm3}: Game playing argument}\label{fig:Game2T3}
\end{figure}

In $\game{1}(A)$, $T[x]\neq$ undefined, if earlier $\Rep$ asked hash-indices of $x$. So,  Bad $\gets$ true if $A$ queries an element present in $\calS$. 
\begin{equation}
\Prob{\game{1}(A):\mathrm{Bad} = \mathrm{true}} = \frac{n}{\abs{\univ}}\label{eq:4T3}
\end{equation}

From \Cref{eq:0T3,eq:1T3,eq:2T3}, \Cref{eq:3T3} and \Cref{eq:4T3}

\begin{equation}
\AdvPrivPubHashBB{B}{A} \leq  \frac{q}{2^\mu} + \frac{n}{\abs{\univ}}
\end{equation}

\end{proof}