### 2016-10-12 EC17-submitted/canonical-definition-instances.tex

\section{Instances of canonical set-multiplicity data structure}
We check whether various set-multiplicity data structures are captured in this definition. Fix set $S$ with elements $x_1, \ldots, x_t \in \bool^{\ell_{in}} $.

\heading{Classical Bloom filter in ROM}
 Fix $k,m > 0$. Define $F: \bool^{\ell_{in}} \rightarrow \bool^{m}}$ as the RO instance; it uses lazy sampling to mimic RO.
\begin{enumerate}
\item $x_1, \ldots, x_t \getsr \Rep_1(S)$.
\item $\pubaux \getsr \Rep_2(|S|,  F(x_1), \ldots, F(x_t))$.
\end{enumerate}

\heading{Linear-prf data structure}
Fix $k,m > 0$. Define $F: \bool^n \times \bool^{\ell_{in}} \rightarrow \bool^{\ell_{out}}$
such that the following hold:
\begin{itemize}
\item $\Rep_1$ takes as input a set $S$ and outputs distinct values $x_1, \ldots, x_t \in \bool^{\ell_{in}}$,
where $t$ depends only on~$|S|$.
\item $\Rep_2$ takes as input $|S|$ and $y_1, \ldots, y_t \in \bool^{\ell_{out}}$, and outputs
$\pubaux$.
\item For any $S$, the distribution on $\pubaux$ output by $\Rep(S)$ is identical to the distribution
on $\pubaux$ computing as follows:
\begin{enumerate}
\item $x_1, \ldots, x_t \getsr \Rep_1(S)$.
\item $k \getsr \bool^n$.
\item $\pubaux \getsr \Rep_2(|S|, F_k(x_1), \ldots, F_k(x_t))$.
\end{enumerate}
\end{itemize}