### 2016-12-07 abstract.tex
pass...

We initiate a provable-security treatment of abstract data
structures by formalizing them as syntactic primitives and
considering their correctness and privacy as first-class security
properties.  Concretely, we explore the security of several data
structures that support (approximate) set-membership tests.  Our
notion of correctness captures an adversary's inability to cause a
data structure to err via its allowed set of queries, and our
privacy notion neatly captures what is leaked by the public portion
of a data structure.  As certain widely used set-membership data
structures, such as Bloom filters, do not meet our privacy notion,
we consider a secondary notion of onewayness capturing the inability
to recover any element of a high-entropy set given the public
portion of the data structure. We also compare our notions to recent
work by Naor and Yogev (Crypto 2015), which also considers
Bloom-filter-like data structures, albeit in a much more limited
way.  Our syntax and notions let us consider security under a
variety of usage scenarios, by clearly surfacing the split between
public and private portions of the data structure.

%We see this work as laying the foundation for a number of follow-on
%efforts, for example to study dynamic data structures (e.g., Bloom
%filter variants whose representation changes over time), and the
%security of primitives built on top of abstract data structures.